---
title: "Data Exploration - US Superstore"
author: "Merilyn Thomas"
format: html
editor: visual
---

## Initial Data Exploration

To start, the following steps will be done in a 'static' manner:

- Produce numerical and graphical summaries to investigate the data. These should include:

  - One-Way and Two-Way Contingency Tables

  - Numerical Summaries (Means, Medians, Standard Deviations, etc) for quantitative variables at levels of categorical variables.

  - Create at least six distinct plots (bar chart, side-by-side bar chart, histogram, box plot, scatter plot, faceted version of one of those, etc)

    - At least four of these should display multivariate information via the type of graph, by utilizing coloring, grouping, etc.

    - All plots should have nice labels and titles.

    - Some kind of faceting should be used somewhere.

    - At least one plot should be a plot that we didn't cover in class (say a heatmap or something like that)

### Reading in the data

```{r}
#| warning: FALSE
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readxl)

superstore_data <- read_excel("US Superstore data.xls", sheet = "Orders")
superstore_data
```

### Numerical Summaries

#### One-Way and Two-Way Contingency Tables

Let's look at numerical summaries to investigate the US Superstore data. First, let's investigate the different categories of orders that are placed. The variable used is called `Category` and the orders can be placed from the following product groups: `Furniture`, `Office Supplies`, or `Technology`.

After observing the one-way contingency table below, we can see that a majority of orders are in the Office Supplies Category.

```{r}
table(superstore_data$Category)
```

The extra factor added to this data. Let's analyze the different sections from where the orders are placed. The extra variable used is called `Segment`. Adding in this extra factor allows us to understand purchasing behaviors across different types of buyers. It helps us understand which Segment of of customer buys different product groups.

```{r}
table(superstore_data$Category, superstore_data$Segment)
```

#### Numerical Summaries (Means, Medians, Standard Deviations, etc)

Numerical Summaries (means, medians, sds, etc.) for quantitative variables at levels of categorical variables. The below data shows the summaries for `Sales` (the selling price of each product) and `Profit` (the profit gained on each product), as they relate to each order **category**.

```{r}
# Mean and Median Sales across different categories
superstore_data %>%
  group_by(Category) %>%
  summarize(
    #Sales Summaries
    Mean_Sales = mean(Sales, na.rm = TRUE),
    Median_Sales = median(Sales, na.rm = TRUE),
    SD_Sales = sd(Sales, na.rm = TRUE),
    
    #Profit Summaries
    Mean_Profit = mean(Profit, na.rm = TRUE),
    Median_Profit = median(Profit, na.rm = TRUE),
    SD_Profit = sd(Profit, na.rm = TRUE)
  )
```

On average based on what is observed, `Technology` category has the highest average sales and profit in comparison to the other categories.

### Graphical Summaries

Creating graphical summaries and plots to help visualize our data. We display multivariate information, facet by variables, and end with a unique plot.

```{r}
library(corrplot)
# bar plot to visualize the one-way contigency table. This gives us a graphical representation showing that Office Supplies is the most common product group in which orders are put. 
# Bar Chart (Univariate Categorical)
ggplot(superstore_data, aes(x = Category)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Number of Orders by Category", 
       x = "Category",
       y = "Number of Orders") 

# visual representation of the two-way contigency table. This plot gives multivariate information regarding `Category` and `Segment`. 
# Side-by-side Bar Chart (Multivariate Categorical)
ggplot(superstore_data, aes(x = Category, fill = Segment)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Orders by Category and Segment", x = "Category", 
       y = "Proportion of Orders", 
       fill = "Product Segment")

# A visual graph to nummerical summaries visualizing the Profit, colored by Category.This helps identify outliers and visualize profit margins. This scatter plot shows the selling price of products and the profit gained from them. Because of the extreme outliers, the plot "zooms out" forming a dense cluster near the origin. 
# Scatter Plot (Multivariate - Numeric and Categorical)
ggplot(superstore_data, aes(x = Sales, y = Profit, color = Category)) +
  geom_point(alpha = 0.5) +
  labs(title = "Sales vs. Profit by Product Category",
       x = "Sales",
       y = "Profit")

# See how sale price of categories has changed between 2014- 2018. It helps us look for seasonlity, cyclical spending habits, or long-term pricing trends. By looking at the line plot below, we cannot see any significant trend in prices rising or falling over these years. Dense clustering shows persistent transaction volumes without an obvious upward or downward trend in base prices. 
# Line Plot with Faceting (Time Series by Category)
ggplot(superstore_data, aes(x = `Order Date`, y = Sales, color = Category)) +
  geom_line(linewidth = 0.75) +
  facet_wrap(~ Category) +
  labs(title = "Sales Over Time by Order Category",
       x = "Order Date",
       y = "Sales (selling price)",
       color = "Order Category")

# Box plots are perfect for visualizing the distribution, median, and 
# spread of order sizes across our different customer Segments. It highlights the typical 
# transaction value while clearly isolating outlier purchases
# Box Plot (Multivariate - Numeric across Categorical)
ggplot(superstore_data, aes(x = Segment, y = Sales, fill = Segment)) +
  geom_boxplot() +
  labs(title = "Sales Distribution by Segment", x = "Segment", y = "Sales")


#This plot provides a deep-dive, multi-dimensional view of our data. 
# By plotting Profit vs. Sales, coloring by Category, and faceting by Segment, we can isolate highly specific business trends. For example, it allows us to investigate if Consumer Furniture sales are inherently less profitable than Corporate Furniture sales.
# Scatter Plot with Faceting (Multivariate, colored, and faceted)
ggplot(superstore_data, aes(x = Sales, y = Profit, color = Category)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ Segment) +
  labs(title = "Profit vs Sales by Segment and Category", x = "Sales", y = "Profit")

# A correlation heatmap instantly quantifies the mathematical relationships between all our key numeric metrics. This is incredibly powerful for pricing strategy. 
# it visually confirms the strength of the negative correlation between 'Discount' and 'Profit', showing exactly how much discounting eats into our bottom line.
# Heatmap (Plot type not typically covered in standard intro)
# Correlation matrix of numeric variables
numeric_vars <- superstore_data %>% select(Sales, Quantity, Discount, Profit) %>% na.omit()
corr_matrix <- cor(numeric_vars)

corrplot(corr_matrix, method = "color", type = "upper", 
         addCoef.col = "black", tl.col = "black", tl.srt = 45,
         title = "Correlation Heatmap of Numeric Variables", 
         mar = c(0,0,1,0))
```
