---
title: "Data Exploration - US Superstore"
author: "Merilyn Thomas"
format: html
editor: visual
---

## Initial Data Exploration

To start, the following steps will be done in a 'static' manner:

- Produce numerical and graphical summaries to investigate the data. These should include:

  - One-Way and Two-Way Contingency Tables

  - Numerical Summaries (Means, Medians, Standard Deviations, etc) for quantitative variables at levels of categorical variables.

  - Create at least six distinct plots (bar chart, side-by-side bar chart, histogram, box plot, scatter plot, faceted version of one of those, etc)

    - At least four of these should display multivariate information via the type of graph, by utilizing coloring, grouping, etc.

    - All plots should have nice labels and titles.

    - Some kind of faceting should be used somewhere.

    - At least one plot should be a plot that we didn't cover in class (say a heatmap or something like that)

### Reading in the data

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readxl)

superstore_data <- read_excel("US Superstore data.xls", sheet = "Orders")
superstore_data
```

### Numerical Summaries

#### One-Way and Two-Way Contingency Tables

Let's look at numerical summaries to investigate the US Superstore data. First, let's investigate the different categories of orders that are placed. The variable used is called `Category` and the orders can be placed from the following product groups: `Furniture`, `Office Supplies`, or `Technology`.

After observing the one-way contingency table below, we can see that a majority of orders are in the Office Supplies Category.

```{r}
table(superstore_data$Category)
```

The extra factor added to this data. Let's analyze the different sections from where the orders are placed. The extra variable used is called `Segment`. Adding in this extra factor allows us to understand purchasing behaviors across different types of buyers. It helps us understand which Segment of of customer buys different product groups.

```{r}
table(superstore_data$Category, superstore_data$Segment)
```

#### Numerical Summaries (Means, Medians, Standard Deviations, etc)

Numerical Summaries (means, medians, sds, etc.) for quantitative variables at levels of categorical variables. The below data shows the summaries for `Sales` (the selling price of each product) and `Profit` (the profit gained on each product), as they relate to each order **category**.

```{r}
# Mean and Median Sales across different categories
superstore_data %>%
  group_by(Category) %>%
  summarize(
    #Sales Summaries
    Mean_Sales = mean(Sales, na.rm = TRUE),
    Median_Sales = median(Sales, na.rm = TRUE),
    SD_Sales = sd(Sales, na.rm = TRUE),
    
    #Profit Summaries
    Mean_Profit = mean(Profit, na.rm = TRUE),
    Median_Profit = median(Profit, na.rm = TRUE),
    SD_Profit = sd(Profit, na.rm = TRUE)
  )
```

On average based on what is observed, `Technology` category has the highest average sales and profit in comparison to the other categories.

### Graphical Summaries

## Initial Data Exploration
